---
title: "SabCLab1_MSiA400"
author: "Sabarish Chockalingam"
date: "October 31, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# MSiA 400 - Lab Assignment 1
## Problem 1
```{r}
# install.packages('expm')
library('expm')
# loading text?file
markov <- read.table('markov100.txt', header=FALSE)
#make data into matrix
P = as.matrix(markov)

#creating inital vector, in state 1
a = rep(0,100)
a[1]=1

prob5 = a%*%(P%^%10)

```
#### 1 a) probability  of being in state 5 after 10 transitions from state 1
```{r}
#probability  of being in state 5 after 10 transitions from state 1
prob5[5]
```
```{r}
# creating different start vector 
a123 = rep(0,100)
a123[1:3]=c(1/3,1/3,1/3)

prob10 = a123%*%(P%^%10)
```
#### 1 b) probability  of being in state 10 after 10 transitions from states 1,2 and 3 being at equal probability
```{r}
prob10[10]
```
#### 1 c) Steady state probability of being in state 1
```{r}
# calculating steady-state probability of state 1
#Subtracting Identity matrix from transpose of P
Q <- t(P)-diag(ncol(P))
#Replacing last row with sum(pi) = 1 constraint
Q[nrow(Q),]=rep(1,ncol(Q))
#right hand side of Q*pi
rhs = c(rep(0,ncol(Q)-1),1)
#Solving for Steady state probabilities
Pi = solve(Q)%*%rhs
Pi[1]
```


```{r}
# calculating mean first passage time to 100

# removing rows and columns corresponding to state 100
B = P[-100,-100]

# subtracting B from Identity matrix
Q = diag(nrow(B))-B

# vector of ones
e = rep(1,nrow(B))

# calculating mean first passage times
m = solve(Q)%*%e
```

#### 1 d) Mean First passage time from state 1 to state 100
```{r}
m[1]
```
## Problem 2
#### 2 a) Contructing Traffic Matrix
```{r}
#loading webtraffic.txt
wb = read.table('webtraffic.txt',header=TRUE)

# Total Traffic Matrix
colsumwb = colSums(wb)
Traffic = matrix(colSums(wb),9,9,byrow = TRUE)


# Displaying Traffic Matrix
Traffic

```
#### 2 b) One step probability transition matrix
```{r}
# From the above matrix it can be seen row 9 and column 1 are filled with 0's.

Traffic[9,1] = 1000

# Constructing one step transi?i?n matrix (divide each element of row by its sum)
P = Traffic/rowSums(Traffic)

#Displaying transition matrix
P
```
#### 2 c) Steady State Probability vector Pi
```{r}
# calculating steady-state probability vector
# Subtracting Identity matrix from transp?s? of P
Q = t(P)-diag(ncol(P))
# Replacing last row with sum(pi) = 1 constraint
Q[nrow(Q),]=rep(1,ncol(Q))
# right hand side of Q*pi
rhs = c(rep(0,ncol(Q)-1),1)
# Solving for Steady state probabilities
Pi = solve(Q)%*%rhs

# Displaying vector
Pi
```
#### 2)d) Average time spent on website
```{r}
# calculating mean first passage time to 9 (last page)

# removing rows and columns corresponding to page 9
B = P[-9,-9]

# subtracting B from Identity matrix
Q = diag(nrow(B))-B

# vector of average times
e = c(0.1,2,3,5,5,3,3,2 )

# calculating mean first passage times
m = solve(Q)%*%e

# average time spent on website is mean first time passage from page 1
m[1]
```
#### 2 e) Checking if adding links to page 6,7 helps uniformly spread traffic.
```{r}
#Calculating new ?u?going Traffic %
dtraff23 = Traffic[2,3]*0.3
dtraff24 = Traffic[2,4]*0.2
Traffic2 = Traffic
Traffic2[2,3]=Traffic[2,3]-dtraff23
Traffic2[2,6]=Traffic[2,6]+dtraff23
Traffic2[2,4]=Traffic[2,4]-dtraff24
Traffic2[2,7]=Traffic[2,7]+dtraff24

# Constructing one?s?ep transition matrix (divide each element of row by its sum)
P2 = Traffic2/rowSums(Traffic2)

# calculating steady-state probability vector
# Subtracting Identity matrix from transpose of P
Q2 = t(P2)-diag(ncol(P2))
# Replacing last row with sum(pi) = 1 ?o?straint
Q2[nrow(Q2),]=rep(1,ncol(Q2))
# right hand side of Q*pi
rhs2 = c(rep(0,ncol(Q2)-1),1)
# Solving for Steady state probabilities
Pi2 = solve(Q2)%*%rhs2

# Displaying vector
Pi2

```
##### comparing
```{r}
# comparing
var.test(Pi,Pi2)
```

```{r}
varPi=var(Pi)
varPi2=var(Pi2)
print(paste('Variance of Steady State Vector var(Pi):',varPi))
print(paste('Variance after adding links var(Pi2):',varPi2))
```

